# In-Class Programming Activity, Day 9

##Nikhita Jain

```{r}
Poker_deck = c(outer((10 * 2:14), 1:4, '+'))

suits <- function(Poker_deck){Poker_deck %% 10}

ranks <- function(Poker_deck){Poker_deck %/% 10}

deal <- function(){sample(Poker_deck,5)}
```

Flush- five cards of the same suit

```{r}
is_flush <- function(five_cards){
  counts <- max(table(suits(five_cards)))
  counts==5
}
is_flush(deal())
```

straight- five consecutively numbered cards

```{r}
is_straight <- function(five_cards){
  counts<- ranks(five_cards)
  max(diff(sort(counts))) == 1
  
}
is_straight(deal())
```

straight flush- five cards of the same suit in consecutive order

```{r}
is_straight_flush <- function(five_cards){
  is_flush(five_cards) == "TRUE" & is_straight(five_cards) =="TRUE"
}
is_straight_flush(deal())
```
Royal flush - five cards of the same suit with ranks 10, Jack, queen, king, ace.

```{r}
is_royal_flush <- function(five_cards){
  is_straight_flush(five_cards) == "TRUE" & min(ranks(five_cards)) == 10
 }
is_royal_flush(deal())
```
three of a kind

```{r}
three_of_a_kind <- function(five_cards){
  counts<- max(table(ranks(five_cards)))
  counts == 3
}
three_of_a_kind(deal())
```

four of a kind- four cards of the same value and different suits

```{r}
four_of_a_kind <- function(five_cards){
  counts<- max(table(ranks(five_cards)))
  counts == 4
}
four_of_a_kind(deal())
```
one pair

```{r}
is_pair <- function(five_cards){
  counts<- max(table(ranks(five_cards)))
  counts==2
}
is_pair(deal())
```

full house- three of a kind and one pair

```{r}
is_full_house <- function(five_cards){
  counts<- max(table(ranks(five_cards)))
  count <- min(table(ranks(five_cards)))
  counts==3 & count ==2 
}
is_full_house(deal())
```


two pair
```{r}
two_pair <- function(five_cards){
  counts <- table(ranks(five_cards))==2
  sum(counts)==2
}
two_pair(deal())
```
```{r}
before_draw<- function(five_cards){
  if (is_royal_flush(five_cards)) return ("royal flush")  
  if (is_flush(five_cards)) return("flush")  
  if (is_straight_flush(five_cards)) return("straight flush") 
  if(is_straight(five_cards)) return("straight")
  if(is_full_house(five_cards)) return("full house")
  if(four_of_a_kind(five_cards)) return("four of a kind")
  if(three_of_a_kind(five_cards)) retunr("three of a kind")
  if(two_pair(five_cards)) return ("two pair")
  if(is_pair(five_cards)) return ("pair")
  return("crappy cards")
}
before_draw(deal())
```

## Test statements

```{r}
scoreActivity::score253(9)
```