# Topic 1 Exercises
#Discussion questions
2.4.1, 2.4.3, 2.4.6:

2.4.1:
a) A flexible approach would be better as it would fit the model better and reduce bias. A large sample size would benefit with a flexible model as opposed to an inflexible one. 

b) A flexible approach would be worse as it would overfit the data as the sample size is small. 

c) Flexible would be better as there would be a greater degree of freedom and a flexible model would obtain a better fit. 

d) A flexible model would be worse as flexibility would just increase the amount of variance.

2.4.3:

a) ![](name of pic)

2.4.6:

A parametric approach makes an assumption for the functional form or shape of f. It reduces the problem of estimating f down to one of estimating the parameters. Non parametric models do not make an assumption for the functional form or shape of f. It requires a very large number of observations to accurately estimate f. 

The advantages of parametric to regression or classification is that not many observations are needed as compared to a non parametric approach and it simplifies it to just estimating the parameters for a model for f. 

The disadvantages are that the model chosen for f could be wrong which would inaccurately estimate f or overfit the observations if more flexible models are used. 

#Computing Assignment
2.4.8, 2.4.9

2.4.8:
a) 

```{r}
##download.file("http://www-bcf.usc.edu/~gareth/ISL/College.csv", destfile="College.csv")
```

```{r}
auto_file_name = "/home/local/MAC/njain/Math-253-Assignments ssh/College.csv"
```

```{r}
college = read.csv(auto_file_name,header=T,na.strings ="?")
```
b) 

```{r}
#instead of fix, use view(college) in console 
```

#Theory Assignment 
2.4.2, 2.4.7

2.4.2: 
a) Regression, inference. This is because we are interested in estimating CEO salary based on certain features of the firm. 
n = 500 firms
p = profit, number of employees and industry

b) Classification, prediction. This is because we are trying to classify whether it was a success or failure and predict the result for each product. 
n = 20 similar products previously launched
p = price charged, marketing budget, comp. price, ten other variables

c) Regression, prediction. Trying to find the quantitative output of % change for US dollar. 
n - 52 weeks of 2012 weekly data
p - % change in US market, % change in British market, % change in German market

2.4.7: 

a)
Obs. 1 distance = 3
Obs. 2 distance = 2
Obs. 3 distance = sqrt(10) ~ 3.16
Obs. 4 distance = sqrt(5) ~ 2.236
Obs. 5 distance = sqrt(2) ~ 1.4
Obs. 6 distance = sqrt(3) ~ 1.7 

b) The observation will be green. This is because when K=1, the closest observation is Obs. 5

c) The observation will be Red. When K =3, observations 1, 3 and 5 are closest. 1 and 3 are red while 5 is green. Thus, our prediction would be red. 

d) We would expect K to be small. This is because a small K would be more flexible and fit a non-linear boundary while a large K would try to fit a more linear boundary as it would take many points into consideration. 

